A fully labelled proof system for intuitionistic modal logics

### 2019-09-27 caleidoscope/caleidoscope.tex
necessitation in completeness and details in other sections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\section{Intuitionistic modal logics}\label{sec:intmod}
The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
%
Starting with a set $\mathcal{A}$ of atomic propositions, denoted by lower case letters $a$, $b$, $c$,\ldots, modal formulas, denoted by capital letters $A$, $B$, $C$, \ldots, are constructed from the grammar:
%
$$
A \coloncolonequals
a \mid A \AND A \mid A \OR A \mid \BOT \mid A \IMP A \mid \BOX A \mid \DIA A
$$
%
\todo{change color?}

@@

\begin{theorem}[\cite{fischer-servi:84,plotkin:stirling:86}]\label{thm:plotkin}
	A formula is a theorem of $\IK$ if and only if it is valid in every bi-relational frame.
\end{theorem}

@@

\begin{figure}[!t]
  \begin{center}
	\fbox{
%	\begin{minipage}{.95\textwidth}
	\begin{tabular}{c@{\quad}c}
		$\vlinf{\rn{id}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a} }{}$
		&
		$\vlinf{\llabrn\bot}{}{\B, \Left, \labels{x}{\BOT} \SEQ \Right}{}$
		%\quad
		%$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, \labels{x}{\TOP}}{}$
		\\\\
		$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
		&
		$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, \labels{x}{A \AND B}}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{B}}$
		\\\\
		$\vliinf{\llabrn\OR}{}{\B, \Left, \labels{x}{A \OR B} \SEQ \Right}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{B} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rlabrn\IMP}{\text{\scriptsize $y$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{\B, \futs xy, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$
		}
		\\\\
		\multicolumn{2}{c}{
 		$\vliinf{\llabrn\IMP}{}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \futs xy, \labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{y}{A}}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}$ 
% 		&$\vliinf{\rlabrn\IMP}{\text{\scriptsize $x \le y \in \B$}}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \Left \SEQ \Right, \labels{y}{A}}{\B, \Left, \labels{y}{B} \SEQ \Right}$
 		}
		\\\\
		$\vlinf{\llabrn\BOX}{}{\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, x \le y, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\BOX}{\text{\scriptsize $y, z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, \futs xy, \accs yz, \Left \SEQ \Right, \labels{z}{A}}$
		\\\\
		$\vlinf{\llabrn\DIA}{\text{\scriptsize $y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, \accs xy, \Left, \labels{y}{A} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\DIA}{}{\B, \accs xy, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, \accs xy, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
		\\
		\multicolumn{2}{c}{
		$\mbox{\hbox to .9\linewidth{\dotfill}}$
		}
		\\
		$\vlinf{\rn{refl}}{}{\B, \Left \SEQ \Right}{\B, \futs xx, \Left \SEQ \Right}$
		&
		$\vlinf{\rn{trans}}{}{\B, \futs xy, \futs yz, \Left \SEQ \Right}{\B, \futs xy, \futs yz, \futs xz, \Left \SEQ \Right}$
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rn{F_1}}{\text{\scriptsize $u$ fresh}}{\B, \accs xy, \futs yz, \Left \SEQ \Right}{\B, \accs xy, \futs yz, \futs xu, \accs uz, \Left \SEQ \Right}$
		}
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rn{F_2}}{\text{\scriptsize $u$ fresh}}{\B, \accs xy, \futs xz, \Left \SEQ \Right}{\B, \accs xy, \futs xz, \futs yu, \accs zu, \Left \SEQ \Right }$		
		}
	\end{tabular}		
%\end{minipage}
}		
  \end{center}
  \caption{System $\labIKp$}
	\label{fig:labIKp}
\end{figure}

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation wpould also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A} \IMP B \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, xRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, xRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. This is the reason why in our system we need the
$\le$ in the relational atoms; we need to go to the future of the
current world.  In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that need the
future-relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.

Finally, in the semantics, the two relations $\rel$ and $\le$ are
strongly connected through the two conditions $\rn{F_1}$ and $\rn{F_2}$. These need to be reflected in the proof
system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.
%
However, these two rule create new labels, and in order to be
complete, the system needs the \emph{monotonicity} rule $\llabrn{mon}$, shown on the left below. 
\begin{equation}
  \label{eq:mono}
  \vlinf{\llabrn{mon}}{}{\lseq{\B, x \le x'}{\Left, \labels{x}{A}}\Right}{
    \lseq{\B, x \le x'}{\Left, \labels{x}{A}, \labels{x'}{A}}\Right}
  \hskip6em
  \vlinf{\rlabrn{mon}}{}{\lseq{\B, x\le x'}{\Left}{\Right, \labels{x'}{A}}}{
    \lseq{\B, x \le x'}{\Left}{\Right, \labels{x}{A}, \labels{x'}{A}}}
\end{equation}
Since this rule is a form of contraction, it would cause the same
problems as contraction in a cut elimination proof. Hence, it is
preferable to have a system in which this rule is admissible. This is
the reason why we have monotonicity incorporated in the rules
$\rn{id}$, $\llabrn\IMP$ and $\llabrn\BOX$ in Figure~\ref{fig:labIKp},
instead of using the rules in~\eqref{eq:nomon}. Then, not only
$\llabrn{mon}$ but also its right-hand side version $\rlabrn{mon}$,
shown on the right in~\eqref{eq:mono} above is admissible.

@@

One can prove this proposition in the same way as one usually proves
admissibity of contraction in a sequent calculus, by inductiop on the
height of the derivation, which in fact would yield a stronger result,
namely that $\llabrn{mon}$ and $\rlabrn{mon}$ are \emph{height
  preserving} admissible for $\labIKp$. However, we do not need this
result in this paper, and therefore we leave it as an exercise to the
reader. Nonetheless, we will give a short proof of
Proposition~\ref{prop:mon-adm} at the end of this section.

@@

	\item $\fm{\DIA A}$
%	\begin{smallequation*}
%		\vlinf{\labrn{id_g}}{}{\B, x \le y; \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{}
%		\reducesto
%	\end{smallequation*}
	\begin{smallequation*}
		\vlderivation{
			\vlin{\llabrn\DIA}{}{\B, \futs xy; \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{
				\vlin{\rn{F_2}}{}{\B, \futs xy, \accs xz, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
					\vlin{\rlabrn\DIA}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
						\vlin{\labrn{id_g}}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}, \labels{u}{A}}{
							\vlhy{}
						}
					}
				}
			}
		}
	\end{smallequation*}
        \qedhere
	\end{itemize}
\end{proof}

@@

\begin{proof}[Proof of Proposition~\ref{prop:mon-adm}]
  The rule $\llabrn{mon}$ can be derived using the general identity and cut:
  \begin{smallequation*}
    \vlderivation{
      \vliiin{\labrn{cut}}{}{\lseq{\B, x \le x'}{\Left, \labels{x}{A}}\Right}{
        \vlin{\labrn{id_g}}{}{\lseq{\B, x \le x'}{\Left, \labels{x}{A}}{\Right, \labels{x'}{A}}}{
          \vlhy{}}}{
        \vlhy{\quad}}{
        \vlhy{\lseq{\B, x \le x'}{\Left, \labels{x}{A}, \labels{x'}{A}}\Right}}
    }
  \end{smallequation*}
  And both these rules are admissible by Proposition~\ref{prop:id} and
  Theorem~\ref{thm:cutfree-compl}. The case for $\rlabrn{mon}$ is
  similar.
\end{proof}

@@

				{\Rightarrow \labels{x}{\BOX (A \IMP B) \IMP (\BOX A \IMP \BOX B)}}

@@

					{\futs xy, \labels{y}{\BOX(A \IMP B)} \Rightarrow \labels{y}{\BOX A \IMP \BOX B}}

@@

						{\futs xy, \futs yz, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A} \Rightarrow \labels{z}{\BOX B}}

@@

							{\futs xy, \futs yz, \futs zw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A} \Rightarrow \labels{u}{B}}

@@

								{\futs xy, \futs yz, \futs zw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A} \Rightarrow \labels{u}{B}}

@@

									{\futs xy, \futs yz, \futs zw, \futs yw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A} \Rightarrow \labels{u}{B}}

@@

										{\futs xy, \futs yz, \futs zw, \futs yw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{u}{B}}

@@

											{\futs xy, \futs yz, \futs zw, \futs yw, \futs uu, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{u}{B}}

@@

												{\B, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{u}{B}, \labels{u}{A}}

@@

												{\B, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B}, \labels{u}{B} \Rightarrow \labels{u}{B}}

@@

		{ \Rightarrow \labels{x}{\BOX (A \IMP B) \IMP (\DIA A \IMP \DIA B)}}

@@

			{ \futs xy, \labels{y}{\BOX (A \IMP B)} \Rightarrow \labels{y}{(\DIA A \IMP \DIA B)}}

@@

				{\futs xy, \futs yz, \labels{y}{\BOX (A \IMP B)}, \labels{z}{\DIA A} \Rightarrow \labels{z}{\DIA B}}

@@

					{\futs xy, \futs yz, \accs zu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A} \Rightarrow \labels{z}{\DIA B}}

@@

						{\futs xy, \futs yz, \accs zu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}

@@

							{\futs xy, \futs yz, \accs zu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}

@@

								{\futs xy, \futs yz, \accs zu, \futs uu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}

@@

									{\B, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}, \labels{u}{A}}

@@

									{\B, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B}, \labels{u}{B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}

@@

			{\Rightarrow \labels{x}{\DIA (A \OR B) \IMP (\DIA A \OR \DIA B)}}

@@

				{\futs xy, \labels{y}{\DIA (A \OR B)} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vliin{\llabrn{\OR}}{}{\futs xy, \accs yz, \labels{z}{A \OR B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}{\vlin {\rlabrn{\OR}}

@@

						{\futs xy, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}

@@

							{\futs xy, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}

@@

								{\futs xy, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}

@@

									{\futs xy, \futs zz, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}

@@

						{\futs xy, \accs yz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}

@@

							{\futs xy, \accs yz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}

@@

		{\Rightarrow \labels{x}{\DIA \BOT \IMP \BOT}}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\section{Intuitionistic modal logics}\label{sec:intmod}
The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
%
Starting with a set $\mathcal{A}$ of atomic propositions, denoted by lower case letters $\fm a$, $\fm b$, $\fm c$,\ldots, modal formulas, denoted by capital letters $\fm A$, $\fm B$, $\fm C$, \ldots, are constructed from the grammar:
%
$$
\fm A \coloncolonequals
\fm a \mid \fm{A \AND A} \mid \fm{A \OR A} \mid \fm \BOT \mid \fm{A \IMP A} \mid \fm{\BOX A} \mid \fm{\DIA A}
$$
%

@@

\begin{theorem}[\cite{fischer-servi:84,plotkin:stirling:86}]\label{thm:plotkin}
	A formula $\fm A$ is a theorem of $\IK$ if and only if $\fm A$ is valid in every bi-relational frame.
\end{theorem}

@@

\begin{figure}[!t]
  \begin{center}
	\fbox{
%	\begin{minipage}{.95\textwidth}
	\begin{tabular}{c@{\quad}c}
		$\vlinf{\rn{id}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a} }{}$
		&
		$\vlinf{\llabrn\bot}{}{\B, \Left, \labels{x}{\BOT} \SEQ \Right}{}$
		%\quad
		%$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, \labels{x}{\TOP}}{}$
		\\\\
		$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
		&
		$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, \labels{x}{A \AND B}}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{B}}$
		\\\\
		$\vliinf{\llabrn\OR}{}{\B, \Left, \labels{x}{A \OR B} \SEQ \Right}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{B} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rlabrn\IMP}{\lb y \mbox{ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{\B, \futs xy, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$
		}
		\\\\
		\multicolumn{2}{c}{
 		$\vliinf{\llabrn\IMP}{}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \futs xy, \labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{y}{A}}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}$ 
% 		&$\vliinf{\rlabrn\IMP}{\text{\scriptsize $x \le y \in \B$}}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \Left \SEQ \Right, \labels{y}{A}}{\B, \Left, \labels{y}{B} \SEQ \Right}$
 		}
		\\\\
		$\vlinf{\llabrn\BOX}{}{\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, x \le y, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\BOX}{\lb y, \lb z \mbox{ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, \futs xy, \accs yz, \Left \SEQ \Right, \labels{z}{A}}$
		\\\\
		$\vlinf{\llabrn\DIA}{\lb y \mbox{ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, \accs xy, \Left, \labels{y}{A} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\DIA}{}{\B, \accs xy, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, \accs xy, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
		\\
		\multicolumn{2}{c}{
		$\mbox{\hbox to .9\linewidth{\dotfill}}$
		}
		\\
		$\vlinf{\rn{refl}}{}{\B, \Left \SEQ \Right}{\B, \futs xx, \Left \SEQ \Right}$
		&
		$\vlinf{\rn{trans}}{}{\B, \futs xy, \futs yz, \Left \SEQ \Right}{\B, \futs xy, \futs yz, \futs xz, \Left \SEQ \Right}$
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rn{F_1}}{\lb u \mbox{ fresh}}{\B, \accs xy, \futs yz, \Left \SEQ \Right}{\B, \accs xy, \futs yz, \futs xu, \accs uz, \Left \SEQ \Right}$
		}
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rn{F_2}}{\lb u \mbox{ fresh}}{\B, \accs xy, \futs xz, \Left \SEQ \Right}{\B, \accs xy, \futs xz, \futs yu, \accs zu, \Left \SEQ \Right }$		
		}
	\end{tabular}		
%\end{minipage}
}		
  \end{center}
  \caption{System $\labIKp$}
	\label{fig:labIKp}
\end{figure}

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation wpould also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. This is the reason why in our system we need the
$\le$ in the relational atoms; we need to go to the future of the
current world.  In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that need the
future-relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.

Finally, in the semantics, the two relations $\rel$ and $\le$ are
strongly connected through the two conditions $\rn{F_1}$ and $\rn{F_2}$. These need to be reflected in the proof
system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.
%
However, these two rule create new labels, and in order to be
complete, the system needs the \emph{monotonicity} rule $\llabrn{mon}$, shown on the left below. 
\begin{equation}
  \label{eq:mono}
  \vlinf{\llabrn{mon}}{}{\lseq{\B, \futs{x}{x'}}{\Left, \labels{x}{A}}\Right}{
    \lseq{\B, \futs{x}{x'}}{\Left, \labels{x}{A}, \labels{x'}{A}}\Right}
  \hskip6em
  \vlinf{\rlabrn{mon}}{}{\lseq{\B, \futs{x}{x'}}{\Left}{\Right, \labels{x'}{A}}}{
    \lseq{\B, \futs{x}{x'}}{\Left}{\Right, \labels{x}{A}, \labels{x'}{A}}}
\end{equation}
Since this rule is a form of contraction, it would cause the same
problems as contraction in a cut elimination proof. Hence, it is
preferable to have a system in which this rule is admissible. This is
the reason why we have monotonicity incorporated in the rules
$\rn{id}$, $\llabrn\IMP$ and $\llabrn\BOX$ in Figure~\ref{fig:labIKp},
instead of using the rules in~\eqref{eq:nomon}. Then, not only
$\llabrn{mon}$ but also its right-hand side version $\rlabrn{mon}$,
shown on the right in~\eqref{eq:mono} above is admissible.

@@

One can prove this proposition in the same way as one usually proves
admissibity of contraction in a sequent calculus, by induction on the
height of the derivation, which in fact would yield a stronger result,
namely that $\llabrn{mon}$ and $\rlabrn{mon}$ are \emph{height
  preserving} admissible for $\labIKp$. However, we do not need this
result in this paper, and therefore we leave it as an exercise to the
reader. Nonetheless, we will give a short proof of
Proposition~\ref{prop:mon-adm} at the end of this section.

@@

	\item $\fm{\DIA A}$
%	\begin{smallequation*}
%		\vlinf{\labrn{id_g}}{}{\B, x \le y; \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{}
%		\reducesto
%	\end{smallequation*}
	\begin{smallequation*}
		\vlderivation{
			\vlin{\llabrn\DIA}{}{\B, \futs xy; \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{
				\vlin{\color{red}\rn{F_2}}{}{\B, \futs xy, \accs xz, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
					\vlin{\rlabrn\DIA}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
						\vlin{\labrn{id_g}}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}, \labels{u}{A}}{
							\vlhy{}
						}
					}
				}
			}
		}
	\end{smallequation*}
        \qedhere
	\end{itemize}
\end{proof}

@@

\begin{proof}[Proof of Proposition~\ref{prop:mon-adm}]
  The rule $\llabrn{mon}$ can be derived using the general identity and cut:
  \begin{smallequation*}
    \vlderivation{
      \vliiin{\labrn{cut}}{}{\lseq{\B, \futs{x}{x'}}{\Left, \labels{x}{A}}\Right}{
        \vlin{\labrn{id_g}}{}{\lseq{\B, \futs{x}{x'}}{\Left, \labels{x}{A}}{\Right, \labels{x'}{A}}}{
          \vlhy{}}}{
        \vlhy{\quad}}{
        \vlhy{\lseq{\B, \futs{x}{x'}}{\Left, \labels{x}{A}, \labels{x'}{A}}\Right}}
    }
  \end{smallequation*}
  And both these rules are admissible by Proposition~\ref{prop:id} and
  Theorem~\ref{thm:cutfree-compl}. The case for $\rlabrn{mon}$ is
  similar.
\end{proof}

@@

				{\SEQ \labels{x}{\BOX (A \IMP B) \IMP (\BOX A \IMP \BOX B)}}

@@

					{\lseq{\futs xy}{\labels{y}{\BOX(A \IMP B)}}{\labels{y}{\BOX A \IMP \BOX B}}}

@@

						{\lseq{\futs xy, \futs yz}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}}{\labels{z}{\BOX B}}}

@@

							{\lseq{\futs xy, \futs yz, \futs zw, \accs wu}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}}{\labels{u}{B}}}

@@

								{\lseq{\futs xy, \futs yz, \futs zw, \accs wu}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}}{\labels{u}{B}}}

@@

									{\lseq{\futs xy, \futs yz, \futs zw, \futs yw, \accs wu}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}}{\labels{u}{B}}}

@@

										{\lseq{\futs xy, \futs yz, \futs zw, \futs yw, \accs wu}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B}}{\labels{u}{B}}}

@@

											{\lseq{\futs xy, \futs yz, \futs zw, \futs yw, \futs uu, \accs wu}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B}}{\labels{u}{B}}}

@@

												{\lseq{\B}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B}}{\labels{u}{B}, \labels{u}{A}}}

@@

												{\lseq{\B}{\labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B}, \labels{u}{B}}{\labels{u}{B}}}

@@

		{ \SEQ \labels{x}{\BOX (A \IMP B) \IMP (\DIA A \IMP \DIA B)}}

@@

			{\lseq{\futs xy}{\labels{y}{\BOX (A \IMP B)}}{\labels{y}{(\DIA A \IMP \DIA B)}}}

@@

				{\lseq{\futs xy, \futs yz}{\labels{y}{\BOX (A \IMP B)}, \labels{z}{\DIA A}}{\labels{z}{\DIA B}}}

@@

					{\lseq{\futs xy, \futs yz, \accs zu}{\labels{y}{\BOX (A \IMP B)}, \labels{u}{A}}{\labels{z}{\DIA B}}}

@@

						{\lseq{\futs xy, \futs yz, \accs zu}{\labels{y}{\BOX (A \IMP B)}, \labels{u}{A}}{\labels{z}{\DIA B}, \labels{u}{B}}}

@@

							{\lseq{\futs xy, \futs yz, \accs zu}{\labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B}}{\labels{z}{\DIA B}, \labels{u}{B}}}

@@

								{\lseq{\futs xy, \futs yz, \accs zu, \futs uu}{\labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B}}{\labels{z}{\DIA B}, \labels{u}{B}}}

@@

									{\lseq{\B}{\labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B}}{\labels{z}{\DIA B}, \labels{u}{B}, \labels{u}{A}}}

@@

									{\lseq{\B}{\labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B}, \labels{u}{B}}{\labels{z}{\DIA B}, \labels{u}{B}}}

@@

			{\SEQ \labels{x}{\DIA (A \OR B) \IMP (\DIA A \OR \DIA B)}}

@@

				{\lseq{\futs xy}{\labels{y}{\DIA (A \OR B)}}{\labels{y}{\DIA A \OR \DIA B}}}
				{\vliin{\llabrn{\OR}}{}{\lseq{\futs xy, \accs yz}{\labels{z}{A \OR B}}{\labels{y}{\DIA A \OR \DIA B}}}{\vlin {\rlabrn{\OR}}

@@

						{\lseq{\futs xy, \accs yz}{\labels{z}{A}}{\labels{y}{\DIA A \OR \DIA B}}}

@@

							{\lseq{\futs xy, \accs yz}{\labels{z}{A}}{\labels{y}{\DIA A}, \labels{y}{\DIA B}}}

@@

								{\lseq{\futs xy, \accs yz}{\labels{z}{A}}{\labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}}

@@

									{\lseq{\futs xy, \futs zz, \accs yz}{\labels{z}{A}}{\labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}}

@@

						{\lseq{\futs xy, \accs yz}{\labels{z}{B}}{\labels{y}{\DIA A \OR \DIA B}}}

@@

							{\lseq{\futs xy, \accs yz}{\labels{z}{B}}{\labels{y}{\DIA A}, \labels{y}{\DIA B}}}

@@

		{\SEQ \labels{x}{\DIA \BOT \IMP \BOT}}
